Supplementary Materials of Impact of IPTp Regimen on Pregnancy Outcomes

### 2025-05-05 code/eda-code/eda_manuscript.qmd
update

*Data Import*
```{r}
# Data Import & Initial Inspection 
# load the Dataset
PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))

promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
```

@@

*Summary Statistics*
Numeric Variables
```{r}
# Summary of numeric variables
promo_data_clean %>% 
  select(where(is.numeric)) %>% 
  summary()

@@

*Categorical Variables*
```{r}
# Frequency tables for categorical variables
promo_data_clean %>% 
  select(where(is.character)) %>% 
  map(~table(.))

@@

*Univariate Plots*
Numeric Variables
```{r}
# Histograms of numeric variables
promo_data_clean %>% 
  select(where(is.numeric)) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  facet_wrap(~key, scales = "free") +
  theme_minimal()

@@

*Categorical Variables*
```{r}
# Bar plots of categorical variables
promo_data_clean %>% 
  select(where(is.character)) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_bar(fill = "lightblue", color = "black") +
  facet_wrap(~key, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

>>>>

*Data Import*

```{r summary-numerics, message=FALSE, warning=FALSE}
library(tidyverse)

# 1) import your cleaned data as a tibble
promo_data_clean <- read_csv(
  here("data", "clean", "PROMO_Data_clean.csv"),
  show_col_types = FALSE
)

# 2) now explicitly call dplyr::select() so there’s no ambiguity
promo_data_clean %>%
  dplyr::select(where(is.numeric)) %>%
  summary()

@@

*Summary Statistics*
Numeric Variables
```{r}
# make sure dplyr is loaded
library(dplyr)

# explicitly call dplyr::select()
promo_data_clean %>%
  dplyr::select(where(is.numeric)) %>%
  summary()

@@

*Categorical Variables*
```{r}
library(purrr)

promo_data_clean %>%
  dplyr::select(where(is.character)) %>%
  map(~ table(.))

@@

*Univariate Plots*
Numeric Variables
```{r}
# 0) load the necessary tidyverse packages
library(tidyverse)   # brings in dplyr, tidyr, ggplot2, purrr, etc.

# 1) read your data (if you haven’t already)
promo_data_clean <- read_csv(
  here("data", "clean", "PROMO_Data_clean.csv"),
  show_col_types = FALSE
)

# 2) select only the numerics, reshape, and plot
promo_data_clean %>%
  dplyr::select(where(is.numeric)) %>%             # force dplyr::select
  tidyr::pivot_longer(                              # modern replacement for gather()
    cols      = everything(),
    names_to  = "variable",
    values_to = "value"
  ) %>%
  ggplot(aes(x = value)) +
    geom_histogram(bins = 30) +
    facet_wrap(~ variable, scales = "free") +
    theme_minimal()

@@

*Categorical Variables*
```{r}
# 0) load the tidyverse (includes dplyr, tidyr, ggplot2, etc.)
library(tidyverse)

# 1) read your cleaned data as a tibble
promo_data_clean <- read_csv(
  here("data", "clean", "PROMO_Data_clean.csv"),
  show_col_types = FALSE
)

# 2) select character columns, pivot longer, and plot
promo_data_clean %>%
  dplyr::select(where(is.character)) %>%        # force dplyr::select
  tidyr::pivot_longer(
    cols      = everything(),
    names_to  = "variable",
    values_to = "value"
  ) %>%
  ggplot(aes(x = value)) +
    geom_bar(fill = "lightblue", color = "black") +
    facet_wrap(~ variable, scales = "free") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2025-05-05 products/manuscript/supplement/Supplementary-Material.qmd
update

# Machine learning & modeling
library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
library(themis)             # Class-imbalance sampling (e.g., SMOTE)
library(dials)              # Parameter tuning grids
library(ranger)             # Random forest engine
library(xgboost)            # Gradient boosting engine
library(generalhoslem)
library(ResourceSelection)
library(rms)

>>>>

# Machine learning & modeling
library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
library(themis)             # Class-imbalance sampling (e.g., SMOTE)
library(dials)              # Parameter tuning grids
library(ranger)             # Random forest engine
library(xgboost)            # Gradient boosting engine
library(generalhoslem)
library(ResourceSelection)
library(rms)
library(tinytex)

### 2025-05-05 products/manuscript/supplement/Supplementary-Material.tex
update

@@

Explain here what each code/file is and does, and in which order (if
any) users need to run thing to reproduce everything. Essentially, give
a full set of instructions to re-generate everything.

\section{Additional Method Details}\label{additional-method-details}

Often, the main manuscript only allows for an overview description of
the methods. Use the supplement to describe all your methods, models and
approaches in a lot of detail. Reference specific parts of your code as
needed.

@@

Show additional results here. Those can be some useful
exploratory/descriptive figures or tables, or results from additional
analyses that didn't make it into the main text.

\emph{Bar Graph of Total Malaria Episodes During Pregnancy by Treatment
Arm}

@@

\emph{Bar Graphs for Gravidity and Parity}

@@

\emph{Parity Distribution by Treatment Arm}

@@

\subsection{Example additional result}\label{example-additional-result}

Table~\ref{tbl-resulttable1} shows an additional table summarizing a
model fit.

Figure~\ref{fig-result2} shows a scatterplot figure produced by one of
the R scripts.

@@

@@

@@

I see that the ROC curve for my model is close to the diagonal, with an
AUC only slightly above 0.5. This tells me that the model doesn't have
strong discriminative ability for predicting adverse outcomes in women
under 25. Additionally, the calibration plot shows that my predicted
probabilities often stray from the ideal diagonal---especially in the
mid-range---indicating that my model's risk estimates don't consistently
match the observed rates. Overall, while gravidity is statistically
significant, my model as a whole isn't very effective at distinguishing
between those who experience adverse outcomes and those who don't, and
its probability estimates need improvement.

@@

\section{References}\label{references}

>>>>

~ ~

@@

The Supplementary Appendix is built with a clear, step‑wise workflow:
raw PROMO trial data are first cleaned and documented, then transformed
into the analysis dataset. Subsequent scripts generate missingness
tables, interaction model diagnostics, machine‑learning tuning results,
and additional stratified analyses. Each RMarkdown file lives in the
code/ subfolders (processing, EDA, analysis), reads from data/, and
writes results into results/. Finally, the Quarto source in
products/manuscript/supplement renders the assembled tables and figures
into Supplementary‑Material.docx. Running the scripts in their numeric
order fully reproduces every table and figure in this appendix.

\subsection{📁 Directory Structure
Overview}\label{directory-structure-overview}

\includegraphics[width=2.39in,height=\textheight,keepaspectratio]{../Images/Project-root.png}

@@

\section{\# Exploratory Analysis: Malaria Episodes, Gravidity/Parity,
and Treatment
Arm}\label{exploratory-analysis-malaria-episodes-gravidityparity-and-treatment-arm}

The four supplementary EDA plots concisely overview key baseline
characteristics by treatment arm and the height--weight relationship at
enrollment. The first panel shows that the majority of women experienced
one malaria episode during pregnancy, with progressively fewer in the
2--3 and ≥4 categories and that counts are consistently higher in the DP
arm than the SP arm across all episode bins. The second panel depicts
gravidity stratified into 1, 2--3, and ≥4 pregnancies, revealing a
similar pattern in both arms but a noticeably smaller proportion of
primigravidae in the SP group. The third panel presents parity
categories (0, 1--2, ≥3), again demonstrating broadly comparable
distributions between arms, with a slight excess of women with ≥3
previous births in the DP arm. Finally, the scatterplot of maternal
height and weight, with a LOESS smoothing curve and 95\% confidence
band, confirms a modest positive association between height and weight
at enrollment, supporting the use of both measures in downstream
modeling.

\emph{Figure S1.Total Malaria Episodes During Pregnancy by Treatment
Arm}

@@

~ ~

\emph{Figure S2: Gravidity Distribution by Treatment Arm}

@@

~ ~

\emph{Figure S3: Parity Distribution by Treatment Arm}

@@

~ ~

\emph{Figure S4:Scatterplot of maternal weight by height with fitted
LOESS smoothing curve (95\% CI) --- exploratory assessment of the
height--weight relationship at enrollment}

@@

~ ~

\subsection{Additional result}\label{additional-result}

\emph{Table 1: Regression Coefficients for the Association Between
Weight and Outcome}

~ ~

@@

\emph{Additional Model Evaluation (Supplement)} Panel A shows the ROC
curve for the gravidity‐only logistic model in women under 25. The curve
lies modestly above the diagonal, reflecting fair discrimination (AUC ≈
0.55). Panel B is the corresponding calibration plot (decile‐based),
which compares observed event rates within probability bins to the ideal
45° line; slight deviations indicate some miscalibration in the highest
and lowest risk groups.

Panel C presents the bootstrap bias‐corrected calibration curve (200
resamples). The ``apparent'' (dotted) and ``bias‐corrected'' (solid)
lines both track closely to the ideal (dashed) reference, confirming
good overall calibration (mean absolute error ≈ 0.016). Panel D displays
the heatmap of cross‐validated AUC across the elastic‐net
penalty--mixture grid, illustrating that intermediate penalty values
(log₁₀λ around --2) and mixture α near 0.8 yielded the best
discrimination.

Finally, Table S7 reports the mean (±SE) cross‐validated ROC AUC for the
three machine‐learning algorithms: LASSO (0.55±0.012), random forest
(0.52±0.013), and XGBoost (0.497±0.014). This confirms that regularized
regression slightly outperformed the tree‐based methods under our tuning
scheme.

\emph{Supplementary Regression Coefficients (Table S8)} The simple
linear model of weight on height yields an intercept of 149.7g (SE 19.8,
p\textless0.001) and a non‐significant slope of 0.23g/cm (SE 0.27,
p=0.43), indicating that while taller women tended to weigh more on
average, this association was weak and did not reach statistical
significance in our sample.

~ ~

\emph{Figure S5: Calibration Plot: Gravidity Model (Age \textless{} 25)}

@@

~ ~

\textbf{Figure S6. Bootstrap‑calibrated calibration curve for the
malaria interaction model}

\includegraphics[width=2.15in,height=\textheight,keepaspectratio]{../Images/rms_model.png}

~ ~

\emph{Figure S7: Calibration curve for malaria interaction model in
women aged \textless25 years}

\includegraphics[width=2.13in,height=\textheight,keepaspectratio]{../Images/CalibrationPlot.png}

~ ~

\emph{Figure S8:Heatmap of five‑fold cross‑validated ROC AUC across the
elastic‑net penalty--mixture grid}

\includegraphics[width=2.14in,height=\textheight,keepaspectratio]{../Images/Heatmap.CV AUC.png}

~ ~

\emph{Table S2:Mean cross‑validated ROC AUC (±SE) for each
machine‑learning mode}

\includegraphics[width=3.88in,height=\textheight,keepaspectratio]{../Images/MeanCV ROC.AUC .png}

@@

The supplementary materials deepen and validate our primary findings by
providing transparent documentation of all data‐processing steps,
extensive methodological details, and additional model diagnostics that
could not be accommodated in the main text. Detailed tables of
missingness, deviance tests, and variance inflation factors reinforce
the robustness of our interaction model. At the same time, the bootstrap
and decile‐based calibration curves confirm its reliability across risk
strata. The elastic‐net heatmap and cross‐validated AUC summaries for
LASSO, random forest, and XGBoost highlight that, under consistent
tuning, regularized regression offers the strongest discrimination in
our dataset. Exploratory bar charts of malaria episodes, gravidity, and
parity by treatment arm contextualize cohort characteristics, and the
gravidity‐only subgroup analyses (ROC and calibration plots) elucidate
how prior pregnancy history predicts adverse outcomes among younger
women. Together, these supplementary analyses enhance confidence in our
conclusions, demonstrate reproducibility, and offer a resource for
readers seeking to replicate or extend this work.
